An approach for the reduction of boundary effects in time-frequency and time-scale representations

### 2020-04-14 Notes/BoundEffRed.tex
notes
new version notes

\documentclass[twoside,letterpaper]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{euscript}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subfig} % or can use subcaption package
\usepackage{color}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{dsfont}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{mathpazo}
\usepackage[left=80pt,right=80pt,top=70pt,bottom=70pt]{geometry}

\title{Tackling boundary effects on kernel-based representations (draft)}
%\author{}
\date{}

@@

\section{Introduction}
In any digital acquisition system, the measured signals are sampled and of finite length. Then, when performing any kernel-based transform, such as synchrosqueezing transform, boundaries effect is going to affect the quality of the representation in the vicinity of the borders. Consequently, signal characteristics, like instantaneous frequencies, appear to be imprecisely determined near borders.

We propose here a method to limit the disruptions caused by the boundary effect. To reduce the influence of boundaries on the kernel-based representations, we proceed in two steps:
\begin{enumerate}
\item Extend the signal by forecasting it.
\item Run the kernel-based algorithm on the extended signal.
\end{enumerate}

\section{Algorithm}
\input{algorithm}

\section{Theoretical performance}
\input{theory}

%\section{Numerical results}
%\input{results}

>>>>

\documentclass[twoside,letterpaper]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{euscript}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subfig} % or can use subcaption package
\usepackage{color}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{dsfont}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{mathpazo}
\usepackage[left=90pt,right=90pt,top=70pt,bottom=70pt]{geometry}

\title{An approach for the reduction of boundary effects in time-frequency and time-scale representations}
%\author{}
\date{}

@@

\section{Introduction}
In any digital acquisition system, the study and the interpretation of the measured signals generally require the use an analysis tool, which enable to point out the useful characteristics of the signal. For example, any  measured physiological signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart. An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure.

In general, the observed quantities are produced by transient phenomena, that can vary rapidly and irregularly. As a consequence, the measured signals exhibit nonstationary behavior. In order to adapt the analysis to nonstationarities, local analysis is generally performed. The short time Fourier transform (STFT) is a typical tool which is build that way, and allows to determine locally the frequency content of a nonstationary signal.

Windowing is a common method for performing local analysis. Among others, STFT, wavelet transform~\cite{Mallat99wavelet}, synchrosqueezing transform (SST), and reassignment~\cite{Auger13time} are representations that fall back on the use of an analysis window. Let $x:I\to\RR$ denote the observed signal, where $I$ denotes the finite interval where the signal is measured. Let $g_s:\RR\to\RR$ denote the analysis window, where $s$ is a shape parameter. The support of $g_s$ is is localized around the origin, and is small with respect to $|I|$. The translation operator is $T_\tau$ defined as:
\[
T_\tau f = f(t-\tau)\ ,\quad \forall\ f:\RR\to\RR\ .
\]
Then, the local analysis of $x$ around the instant $\tau\in I$ rely on the evaluation of the following dot product:
\begin{equation}
V_x(s,\tau) = \langle x, T_\tau g_s \rangle_I \ .
\label{eq:windowing}
\end{equation}
A major shortcoming of this technique occurs when analyzing the signal $x$ near the boundaries of the interval $I$. Indeed, at these points, half of the information is missing. Consequently the results of the dot product~\ref{eq:windowing} is disturbed by this lack of information. This phenomenon is the so-called \emph{boundary effect}. We display, on the left of Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description). The distortion resulting from the boundary effect is clearly visible on both sides of this representation. Indeed, while in the major central part of the image, clear lines stand out, they become blurred as they approach the left and right boundaries of the image. A zoom on the right boundary of this SST, displayed on the right of Fig.~\ref{fig:ex.intro}, emphasizes the result of boundary effect. More generally, signal characteristics, like instantaneous frequencies or amplitudes appear to be imprecisely determinable in the vicinity of the boundaries.

\begin{figure}
\includegraphics[width=.49\textwidth]{SSTintro.eps}
\includegraphics[width=.49\textwidth]{zoomSSTintro.eps}
\caption{Synchrosqueezing transform of a PPG (left) with a zoom on its right boundary (right). }
\label{fig:ex.intro}
\end{figure}

The purpose of this paper is to provide a fast algorithm allowing to tackle boundary effect and limit the disruptions it may cause on the signals representations based on windowing. In that aim, we proceed in two steps:
\begin{enumerate}
\item \emph{Extend the signal by forecasting it.} The aim is to use a dynamic model to predict the values taken by the measured signal outside the measurement interval. Then, once this operation is done, we have access to an extended signal defined on a larger interval $I_\Delta$, where $\Delta$ denotes the size of the extension on both boundaries of $I$. 
\item \emph{Run the local analysis tool on the extended signal.} Assuming that the support of the analysis window is smaller that $2\Delta$, the local analysis near the boundary of $I$ is now possible without lack of information thanks to knowledge brought by the extension on both sides. 
\end{enumerate}
Thus, assuming that the quality of the extension step is sufficient, the representation obtained that way will be less sensitive to the boundary effect than the result of the analysis tool applied directly to the non-extended signal. Notice that, in order to obtain a fast algorithm the extension step we propose must have a low computational cost.

\section{Algorithm}
\label{se:algo}
\input{algorithm}

\section{Theoretical performance}
\label{se:theoretical}
\input{theory}

\section{Numerical results}
\input{results}

\bibliographystyle{plain}
\bibliography{AM041320}

### 2020-04-14 Notes/algorithm.tex
new version notes

\paragraph{Dynamic model and forecasting.} Establishing a dynamic model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\[
\bY = f(\bX)\ .
\] 
We consider here a naive dynamic model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\end{equation}
where $\bA\in\RR^{M\times M}$. This is a linear dynamic model, which can be written equivalently in function of the sub-signals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}

@@

\subsection{Kernel-based representation}

\paragraph{Transform restriction.} 
Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the kernel-based representation we are interested in. It can be, for instance, such as short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS). Here, $F$ typically denotes the size of the representation along the frequency axis. Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges. To limit these phenomena, we apply the representation to the estimated extended signal $\tilde\bx$. This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$. Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given by:
\begin{equation}
\ccF_N^{\mathrm{ext}}(\bx)[f,n] = \ccF_{N+2L}(\tilde \bx)[f,L+n]\ ,\quad\forall f\in\{0,\cdots,F-1\},\ n\in\{0,\cdots,N-1\}\ .
\label{eq:restriction}
\end{equation}
This amounts to restricting the representation $\ccF_{N+2L}(\tilde \bx)$ to the original measurement interval of $\bx$. For the sake of simplicity, we denotes the restriction operator by $\cR$, where $\cR:\RR^{F\times (N+2L)}\to\RR^{F\times N}$. Consequently, we have:
\begin{equation*}
\ccF_N^{\mathrm{ext}}(\bx) = \cR\left( \ccF_{N+2L}(\tilde \bx) \right) \ .
\end{equation*}

\paragraph{Conclusion.}
Finally, the global procedure we implement to reduce boundary effects on kernel-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.

>>>>

\paragraph{Dynamic model and forecasting.} Establishing a dynamic model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\[
\bY = f(\bX)\ .
\] 
We consider here a naive dynamic model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\label{eq:dyn.model}
\end{equation}
where $\bA\in\RR^{M\times M}$. This is a linear dynamic model, which can be written equivalently in function of the sub-signals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}

@@

\subsection{Representation}

\paragraph{Transform restriction.} 
Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the time-frequency or time-scale representation we are interested in. It can be, for instance, such as short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS). Here, $F$ typically denotes the size of the representation along the frequency axis. Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges. To limit these phenomena, we apply the representation to the estimated extended signal $\tilde\bx$. This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$. Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given by:
\begin{equation}
\ccF_N^{\mathrm{ext}}(\bx)[f,n] = \ccF_{N+2L}(\tilde \bx)[f,L+n]\ ,\quad\forall f\in\{0,\cdots,F-1\},\ n\in\{0,\cdots,N-1\}\ .
\label{eq:restriction}
\end{equation}
This amounts to restricting the representation $\ccF_{N+2L}(\tilde \bx)$ to the original measurement interval of $\bx$. For the sake of simplicity, we denotes the restriction operator by $\cR$, where $\cR:\RR^{F\times (N+2L)}\to\RR^{F\times N}$. Consequently, we have:
\begin{equation*}
\ccF_N^{\mathrm{ext}}(\bx) = \cR\left( \ccF_{N+2L}(\tilde \bx) \right) \ .
\end{equation*}

\paragraph{Conclusion.}
Finally, the global procedure we implement to reduce boundary effects on windowing-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.

### 2020-04-14 Notes/results.tex
new version notes

\subsection{Evaluation the forecasting performance}
In that section, we first evaluate the quality of the forecasting step. That one depends on at least two conditions:
\begin{itemize}
\item The validity of the dynamic model we use to implement forecasting.
\item An optimal choice of the set of hyperparameters on which our algorithm depends. 
\end{itemize}
We study that two points in the following.

\subsubsection{Validation of the linear dynamic model}
\label{ssse:validation}
We proved that the linear dynamic model is sufficient to describe the dynamical behavior of signals as long as the adaptive harmonic model is relevant to describe those signals. We confirm here this theoretical result on real-life signals. To this end, we apply Algorithm~\ref{alg:extension} to a larger number of pieces of a respiratory signal. The signal is 50 minutes-long and is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}. From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$.

@@

The forecasting Algorithm~\ref{alg:extension} is applied to each of the signals in the dataset we built. These ones are extended of 7 second on each border, corresponding to $L =700$. Moreover, we choose the following set of hyperparameters: $M=\lfloor 1.5L\rfloor$ and $K=\lfloor2.5M\rfloor$. We justify such a choice in section~\ref{ssse:hyperparam}. To quantify the quality of our forecasting approach, we evaluate the Mean Square Error $\mathrm{MSE}(\tilde\bx)$ of the estimated extended signal, namely:
\[
\mathrm{MSE}(\tilde\bx) = \dfrac1{2L}\|\tilde\bx_{\rm bw} -\bx_{\rm bw}\|^2 + \dfrac1{2L}\|\tilde\bx_{\rm fw} -\bx_{\rm fw}\|^2\ .
\]
The mean and the standard deviation of the MSE with respect to the simulations are given in Table~\ref{tab:mse}.

The results are compared with the strategy consisting in a zero-padding extension of the signal. The results neatly shows on improvement of the quality of the forecasting. Indeed, in average, the forecasting MSE is improved of almost $20\%$. Nevertheless, the variance forecasting MSE shows a huge variability in the quality of the estimation. This is probably due to the outliers and pulse that can be find in the respiratory signal. That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamic model.

@@

\subsubsection{Influence of the hyperparameters}
\label{ssse:hyperparam}
We evaluate the influence of the hyperparameters $K$ and $M$ on the forecasting MSE. This study is performed on a 32 second-long photoplethysmogram (PPG) signal, which is displayed on the top of Figure~\ref{fig:ppg}. This signal is sampled at $\fs=125$~Hz.

@@

In this study, the extension length $L$ is fixed because the second step of the Algorithm~\ref{alg:boundary} determines the length $L$. Indeed, in order to avoid boundary effect, this one should be at least equal to the half of the width of the window used to perform the kernel-based transform. In practice, we choose to extend the signal if 2~seconds on each side, \ie~$L=250$.
\begin{enumerate}[a),leftmargin=*]
\item 
We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals length $M$. Meanwhile, the value of $K$ is unchanged and set to $K=10L$. The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $M$ on the forecasting MSE. 
%When $M$ is smaller than $L$, the forecasting algorithm cannot provide accurate forecasting because . As soon as $M$ is greater than $L$, the MSE will not be influenced by the sub-signal lengths. Indeed, the information provided by
\item
We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals dataset size $K$, when the value of $M$ is set to $M=\lfloor 1.5L\rfloor$. The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $K$ on the forecasting MSE.
\end{enumerate}

\begin{figure}
\includegraphics[width=.5\textwidth]{mseM.eps}
\includegraphics[width=.5\textwidth]{mseK.eps}
\caption{Influence of the size parameters on the forecasting MSE. Left: MSE in function of the sub-signals length $M$. Right: MSE in function of the sub-signal dataset size $K$.}
\label{fig:influence.M}
\end{figure}

\paragraph{Influence of noise.}
The same kind of study is done. The results are displayed on Figure~\ref{fig:influence.noise}. Here, $M=\lfloor 1.5L\rfloor$ and $K=2M$.

\begin{figure}
\centering
\includegraphics[width=.7\textwidth]{mseNoise.eps}
\caption{Influence of the noise variance on the forecasting MSE.}
\label{fig:influence.noise}
\end{figure}

\subsection{Evaluation of the quality of the boundary effect reduction}
The quality of the boundary effect reduction step must be evaluated of the synchrosqueezing representation. To that aim, we compare the obtained synchrosqueezing transform to the optimal synchrosqueezing transform $\ccF_N^\mathrm{opt}(\bx)$. The optimal synchrosqueezing transform is defined as the restriction of the synchrosqueezing of the ground truth extended signal $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$. Therefore:
\begin{equation*}
\ccF_N^\mathrm{opt}(\bx) = \cR\left( \ccF_{N+2L}(\bx^\mathrm{ext}) \right) \ .
\end{equation*}

This distance is evaluated on the dataset of respiratory signals introduced in section~\ref{ssse:validation}. The results are summarize in Table~\ref{tab:otd}. The OTD averaged over the simulations clearly shows that $\ccF_\mathrm{FC}$ outperforms $\ccF$. This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate synchrosqueezing transform. Moreover, the OTD decreases by almost $50\%$ from the standard SST $\ccF$ to the improved SST $\ccF_{\mathrm{FC}}$. Thus, the improvement obtained by forecasting in the time domain is amplified in the time-frequency domain.

\begin{table}
\centering
\caption{Performance of the Boundary effect reduction on the SST.}
\begin{tabular}{|c||c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{OTD to $\ccF_N^{\mathrm{opt}}(\bx)$} \\
   \cline{2-3}
      & Mean & Standard deviation\\
   \hhline{|=#=|=|}
   $\ccF_N^{\mathrm{ext}}(\bx)$ & $4.705\times 10^{-3}$ & $3.147\times 10^{-3}$ \\
   \hline
   $\ccF_N(\bx)$ & $9.195\times 10^{-3}$ & $3.453\times 10^{-3}$ \\
   \hline
\end{tabular}
\label{tab:otd}
\end{table}

>>>>

\subsection{Evaluation the forecasting performance}
In that section, we first evaluate the quality of the forecasting step. That one depends on at least two parameters:
\begin{itemize}
\item The noise variance $\sigma^2$.
\item The size of the training dataset $K$. 
\end{itemize}
In subsections~\ref{ssse:res.sine} and~\ref{ssse:res.ahm}, we study the influence of these parameters. A comparison with the theoretical results of section~\ref{se:theoretical} is also available.

\subsubsection{Sum of sine waves}
\label{ssse:res.sine}
We proved that the linear dynamic model is sufficient to catch the dynamical behavior of signals taking the form~\eqref{eq:sum.sine}. In order to validate this theoretical result, we apply the forecasting Algorithm~\ref{alg:extension} to a large number of realizations of the random vector $\bz$ such that:
\[
\bz = \bx+ \sigma\bw\ ,
\]
where
\[
\bx[n] =\cos\left(2\pi p_1 \dfrac{n}{M} \right) + R\cos\left(2\pi p_2 \dfrac{n}{M} \right)\ ,\quad\forall\ n\in\{1,\ldots,N\}\ ,
\]
and $N=10^4$, $M=750$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.

\paragraph{Influence of the noise variance $\sigma^2$.} Here, the size of the training dataset is set to $K=1875$. Then, the forecasting algorithm is run on 100 realizations on the discrete signal $\bx$ for 100 different values of $\sigma^2$  logarithmically spaced from $5.10^{-3}$ to $10^{-1}$. For each of these values, we determine the experimental bias $\mu_{xp}[\ell]$ and variance $\gamma_{xp}[\ell,\ell]$ in function of the forecasting sample index $\ell$ (going from 1 to $L=500$). These results are displayed on Figure~\ref{fig:res.noise.sine}. Each color corresponds to these experimental results obtained for a given value of $\sigma$. On the left of Figure~\ref{fig:res.noise.sine}, the result show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$, and always insignificant with respect to the signal values. On the right of Figure~\ref{fig:res.noise.sine}, we display the normalized experimental variance, that is $\frac{\gamma[\ell,\ell]}{\sigma^2}$. This highligths the fact that the forecasting variance increases linearly with respect to $\sigma^2$. More surprisingly, this result shows that the forecasting variance decreases with $\ell$, what is counterintuitive.  
\begin{figure}
\includegraphics[width=.49\textwidth]{biasNoiseSine.eps}
\includegraphics[width=.49\textwidth]{VarianceNoiseSine.eps}
\caption{Evolution of the experimental forecasting bias (left) and the normalized forecasting variance (right) in function of the forecasting sample index. Each color characterizes the result for a given value of $\sigma$.}
\label{fig:res.noise.sine}
\end{figure}

\paragraph{Influence of the training dataset size $K$.} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$. Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for 25 different values of $K$, logarithmically spaced from $3.5\times 10^{3}$ to $6\times 10^{3}$. For each of these values, we determine the experimental bias $\mu_{xp}[\ell]$ and variance $\gamma_{xp}[\ell,\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$). These results are displayed on Figure~\ref{fig:res.size.sine}. Each color corresponds to these experimental results obtained for a given value of $K$. On the left of Figure~\ref{fig:res.size.sine}, as expected, the experimental bias vanishes. On the right of Figure~\ref{fig:res.size.sine}, we display the product of $K$ with the normalized experimental variance, that is $\frac{\gamma[\ell,\ell]}{\sigma^2}$. This result shows the asymptotic behavior of the variance. Indeed, as soon as $K$ is sufficiently high, the product $K\gamma[\ell,\ell]$ is approximately independent of $K$.
\begin{figure}
\includegraphics[width=.49\textwidth]{BiasKSine.eps}
\includegraphics[width=.49\textwidth]{VarianceKSine.eps}
\caption{Evolution of the experimental forecasting bias (left) and the normalized forecasting variance (right) in function of the forecasting sample index. Each color characterizes the result for a given value of $K$.}
\label{fig:res.size.sine}
\end{figure}

\paragraph{Conclusion.}
Both previous results allows us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, which is empirically summarized as follows:
\begin{equation}
\gamma[\ell,\ell] \sim \dfrac{\sigma^2}{K}g[\ell] \ ,\quad\mbox{when}\ \sigma^2\to 0\ \mbox{and}\ K\to\infty \ ,
\end{equation} 
where $g$ is a bounded positive function.

\subsubsection{Adaptive harmonic model}
\label{ssse:res.ahm}
We now consider a signal whose instantaneous frequencies and amplitudes of its components vary over time. The toy signal we use takes the following form:
\[
\bx[n] = \cos\left(2\pi \phi_1[n] \right) + R[n]\cos\left(2\pi \phi_2[n] \right)\ ,\quad\forall\ n\in\{1,\ldots,N\} \ ,
\] 
where the instantaneous amplitude $R$ is given by:
\[
R[n] = 1.4 + 0.2\cos\left(4\pi\frac{n}{N}\right)\ ,
\]
and the instantaneous phases are such that:
\begin{align*}
\phi_1[n] &= \frac{p_1}{M}\left( n + \frac{0.01}{2\pi}\cos\left(2\pi\frac{n}{N}\right) \right) \\
\phi_2[n] & = p_2\frac{n}{M} + \frac{20}{2N\fs}n^2
\end{align*}
Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$. Numerically, we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$ and $R=1.4$.

To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals. These methods are:
\begin{itemize}
\item The \emph{Extended Dynamic Mode Decomposition (EDMD)} has been developped by Williams \etal~\cite{Williams15data}. The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows to catch dynamic of nonlinear systems~\cite{Korda18linear}.
\item The \emph{Gaussian Process Regression (GPR)}~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model. That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}. 
\end{itemize}

To quantify the global quality (\ie~not depending on $\ell$) of the forecasting approaches, we evaluate the Experimental Mean Square Error $\mathrm{MSE_{xp}}(\tilde\bx)$ of the forward forecast extended signals, namely:
\begin{align}
\label{eq:mse}
\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\
\nonumber
&= \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[\ell]^2 + \bgamma_{\xp}[\ell,\ell]\ .
\end{align}
where $\bx^\mathrm{ext}$ is the ground-truth extended signal, that is: $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$. Then, as long as the bias $\bmu[\ell]$ and the variance $\bgamma[\ell,\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either. Corresponding results are given in Table~\ref{tab:mse.sine}. They show that the naive extension we propose gives satisfying results, even though the oher methods, more sophisticated gives MSE values that are somewhat smaller. Netherveless, a major limit of those methods is the computing time they require, which prevent them from being used to exploit real-time data. Thus, {\sf SigExt} is the method that optimize the trade-off between the forecasting quality and the computing time as extension method. That is why, it is implemented in our algorithm for the reduction of boundary effect.

\begin{table}
\centering
\caption{Performance of the extension methods.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{*}{Computing time (sec.)} \\
   \cline{2-3} & Mean & Standard deviation & \\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.15$ \\
   \hline
   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.53$\\
   \hline
   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.33$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

\subsection{Evaluation of the quality of the boundary effect reduction}

\subsubsection{Metrics}
The quality of the boundary effect reduction must be evaluated on the synchrosqueezing representation. To that aim, we compare the obtained synchrosqueezing transform to the optimal synchrosqueezing transform $\ccF_N^\mathrm{opt}(\bx)$. The optimal synchrosqueezing transform is defined as the restriction of the synchrosqueezing of the ground-truth extended signal $\bx^\mathrm{ext}$. Therefore, we have:
\begin{equation*}
\ccF_N^\mathrm{opt}(\bx) = \cR\left( \ccF_{N+2L}(\bx^\mathrm{ext}) \right) \ .
\end{equation*}

\subsubsection{Respiratory signal}

The signal is 50 minutes-long and is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}.

@@

From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$. On each of these pieces of signal, we implement the Algorithm~\ref{alg:boundary} on diverse time-frequency and time-scale representations. The corresponding results are given in Table~

The forecasting Algorithm~\ref{alg:extension} is applied to each of the signals in the dataset we built. These ones are extended of 7 second on each border, corresponding to $L =700$. Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$. As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$. The average and the standard deviation of the MSE~\ref{eq:mse} with respect to the simulations are given in Table~\ref{tab:mse}.

@@

The results are compared with the strategy consisting in a zero-padding extension of the signal. These results  show a significant improvement of the quality of the forecasting. Indeed, in average, the forecasting MSE is improved of almost $20\%$. Nevertheless, the variance forecasting MSE shows a huge variability in the quality of the estimation. This is probably due to the outliers and pulse that can be find in the respiratory signal. That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamic model.

Even though the performance of the forecasting algorithm is moderate, the boundary effect can be reduced dramatically on the time-frequency and time-scale representations. We evaluate the OTD to the optimal representation for different time-frequency and time-scale representation. Notice that the extension length $L$ is set accordingly to the window length used by the time-frequency analysis tool. For instance, here the window length we use to evaluate the STFT is of 1500 samples. To prevent the STFT from being sensitive to the boundary effect, we set $L=750$. In this way, when evaluating the spectral content of the signal near its boundaries, the analysis is not limited by a lack of information all along the window support. From now on, all results are given for $L$ at equal to the half of the width of the window used in the time-frequency transform.

\begin{table}
\centering
\caption{Performance of the Boundary effect reduction on TFR.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Time-Frequency Representation} \\
   \cline{2-4}
      & STFT & SST & RS\\
   \hhline{|=#=|=|=|}
   Without extension & $2.16\times 10^{-2}$ & $5.26\times 10^{-3}$ & $3.07\times 10^{-2}$ \\
   \hline
   {\sf SigExt} & $1.72\times 10^{-2}$ & $4.00\times 10^{-3}$ & $2.43\times 10^{-2}$ \\
   \hline
   EDMD & $1.75\times 10^{-2}$ & $4.23\times 10^{-3}$ & $2.45\times 10^{-2}$ \\
   \hline
   GPR & $1.84\times 10^{-2}$ & $4.10\times 10^{-3}$ & $2.48\times 10^{-2}$ \\
   \hline
\end{tabular}
\label{tab:otd.tho}
\end{table}

\subsubsection{Photoplethysmogram}
\label{ssse:ppg}
We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal, sampled at $\fs=125$~Hz. A 32 second-long piece of this signal is displayed on the top of Figure~\ref{fig:ppg}. The estimated 2-second extension on both sides of this signal is superimposed to the ground-truth signal is provided on the bottom of Figure~\ref{fig:ppg}.

@@

\begin{figure}
\includegraphics[width=.49\textwidth]{SSTBoundEffRed.eps}
\includegraphics[width=.49\textwidth]{zoomSSTBoundEffRed.eps}
\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (left) with a zoom on its right boundary (right). }
\label{fig:ppg.boundeffred}
\end{figure}

%\begin{enumerate}[a),leftmargin=*]
%\item 
%We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals length $M$. Meanwhile, the value of $K$ is unchanged and set to $K=10L$. The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $M$ on the forecasting MSE. 
%When $M$ is smaller than $L$, the forecasting algorithm cannot provide accurate forecasting because . As soon as $M$ is greater than $L$, the MSE will not be influenced by the sub-signal lengths. Indeed, the information provided by
%\item
%We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals dataset size $K$, when the value of $M$ is set to $M=\lfloor 1.5L\rfloor$. The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $K$ on the forecasting MSE.
%\end{enumerate} 
%\begin{figure}
%\includegraphics[width=.5\textwidth]{mseM.eps}
%\includegraphics[width=.5\textwidth]{mseK.eps}
%\caption{Influence of the size parameters on the forecasting MSE. Left: MSE in function of the sub-signals length $M$. Right: MSE in function of the sub-signal dataset size $K$.}
%\label{fig:influence.M}
%\end{figure}

We divide the signal into 32-second long pieces, and apply Algorithm~\ref{alg:boundary} on each piece. We provide in Table~\ref{tab:otd.ppg} the OTD to the optimal time-frequency representation averaged over the signals. For all the considered time-frequency representations, the results clearly shows that our algorithm reduce the influence of the boundary effect. This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate representations.

On Fig.~\ref{fig:ppg.boundeffred}, we display the SST resulting from {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}. We clearly observe an improvement of the quality of the STT near boundaries. Indeed, the blurring visible on the zoom on the right boundary of the SST has almost vanished (see right of Fig.~\ref{fig:ppg.boundeffred}). The real-time tracking of the instantaneous frequencies contained the measured signal is therefore largely facilitated.

\begin{table}
\centering
\caption{Performance of the Boundary effect reduction on TFR.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Time-Frequency Representation} \\
   \cline{2-4}
      & STFT & SST & RS\\
   \hhline{|=#=|=|=|}
   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
   \hline
   {\sf SigExt} & $1.22\times 10^{-3}$ & $7.31\times 10^{-2}$ & $1.11\times 10^{-1}$ \\
   \hline
   EDMD & $9.83\times 10^{-4}$ & $5.59\times 10^{-2}$ & $9.80\times 10^{-2}$ \\
   \hline
   GPR & $1.14\times 10^{-3}$ & $7.02\times 10^{-2}$ & $1.07\times 10^{-1}$ \\
   \hline
\end{tabular}
\label{tab:otd.ppg}
\end{table}
